Prediction of 90-day mortality after Total Hip Arthroplasty: a simplified and externally validated model based on observational registry data from Sweden, England and Wales

### 2020-08-27 reports/article.Rmd
Completeness and coverage. Fix #58

The inclusion period started 2008 since the American Society for Anesthesiologists (ASA) class and Body Mass Index (BMI) were systematically recorded in SHAR from then on. The observation period ended 2015, since we had access to co-morbidity from the national patient register up to this point. Deterministic data linkage was performed by 10-digit identity numbers, assigned to all Swedish residents at birth or immigration. The data linkage has been described previously.[@Cnudde2016] Age, sex, BMI, ASA class, type of hospital and year of surgery were collected from the SHAR, with a completeness of 96-98 %.[@Karrholm2018] Data on educational level were recorded for more than 98 % of the population with 85 % accuracy in the longitudinal integration database for health insurance and labor market studies (LISA) from Statistics Sweden.[@Ludvigsson2019] Civil status, were also collected from LISA and co-morbidity during the year preceding index surgery from the National Patient Register (NPR). NPR contains diagnoses coded by ICD-10 from in- and outpatient episodes in all private and public hospitals. Completeness for NPR is above 99 % and 85-95 % of all diagnostic codes are valid.[@Ludvigsson2011] Death dates were linked from the national population register.

@@

Similar inclusion criteria were applied to the validation cohort. We excluded no patients based on missing information on BMI, educational level, or civil status however, since those variables were not used in the final model. `r format(N_njr$tot, big.mark = ",")` patients were included (Figure \@ref(fig:flowchartnjr)).
NJR data was linked to the secondary uses service database from the National Health Service (NHS) with co-morbidity from the hospital episode statistics (HES) database in England, and the patient episode database for Wales.[@Thorn2016] Out-patient deaths were linked from the office of national statistics based on NHS-numbers, as well as patient name, age, sex and address.[@Sayers2016]

>>>>

~~All public and private Swedish hospitals report their THAs to SHAR, yielding a coverage of 100 %, with an individual patient completeness of 98 %.[@Karrholm2018]~~ Our inclusion period started 2008 since the American Society for Anesthesiologists (ASA) class and Body Mass Index (BMI) were systematically recorded in SHAR from then on. The observation period ended 2015, since we had access to co-morbidity from the national patient register up to this point. Deterministic data linkage was performed by 10-digit identity numbers, assigned to all Swedish residents at birth or immigration. The data linkage has been described previously.[@Cnudde2016] Age, sex, BMI, ASA class, type of hospital and year of surgery were collected from the SHAR, with a completeness of 96-98 %.[@Karrholm2018] Data on educational level were recorded for more than 98 % of the population with 85 % accuracy in the longitudinal integration database for health insurance and labor market studies (LISA) from Statistics Sweden.[@Ludvigsson2019] Civil status, were also collected from LISA and co-morbidity during the year preceding index surgery from the National Patient Register (NPR). NPR contains diagnoses coded by ICD-10 from in- and outpatient episodes in all private and public hospitals. Completeness for NPR is above 99 % and 85-95 % of all diagnostic codes are valid.[@Ludvigsson2011] Death dates were linked from the national population register.

@@

Similar inclusion criteria were applied to the validation cohort. We excluded no patients based on missing information on BMI, educational level, or civil status however, since those variables were not used in the final model. `r format(N_njr$tot, big.mark = ",")` patients were included (Figure \@ref(fig:flowchartnjr)).

~~Reporting to NJR is mandatory for all hospitals run by the National health service (NHS) in England and Wales.[@NationalJointRegistryNJR] The register coverage is thus 100&nbsp;% for such hospitals, although privately founded THAs are not included. Patient participation is based on informed consent. Consent rates varies slightly between years and regions but were 92.3&nbsp;% in both England and Wales in 2018.[@TheNJREditorialBoard]~~ NJR data was linked to the secondary uses service database from the National Health Service (NHS) with co-morbidity from the hospital episode statistics (HES) database in England, and the patient episode database for Wales.[@Thorn2016] Out-patient deaths were linked from the office of national statistics based on NHS-numbers, as well as patient name, age, sex and address.[@Sayers2016]
